Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-09 foreground_delay_spectrum_modeling.tex
updated beam description. small edits through section 4.2.1

The power pattern of a MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active research. The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.  To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the detailed simulation. We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.

@@

\subsection{Foreground Model}\label{sec:foreground}

An instrument such as the MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) for imaging purposes. In the context of EoR H{\sc i} power spectrum, it has been known that unsubtracted forground sources anywhere in the visible hemisphere up to the horizon directly contaminate the spatial frequency modes in the power spectrum. In addition, this contamination also spills over into the relatively cleaner regions, called {\it EoR window}, due to spectral properties of the instrument and the foregrounds \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.

\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account theremal noise effects in the {\it Eor window}. \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise. In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum. We use an all-sky foreground emission model that consists of diffuse and bright compact components.

@@

For the diffuse component, we use an all-sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all-sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. Using the same source model, we also obtain the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.

A low resolution version of the diffuse foreground model used is show in figure~\ref{fig:DSM}. Contours of the MWA tile power pattern shown in figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 0.125$) in that direction. In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times lesser.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 0.125$) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

>>>>

The precise power pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation. The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.  To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO simulation. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.

@@

\subsection{Foreground Model}\label{sec:foreground}
%todo
The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon. As we have seen above, foreground sources contaminate the region inside the ``wedge'' with an amplitude that depends on the primary beam amplitude in that direction, with sources at the horizon being closest to the edge of the wedge.

The degree to which these sources leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the edge of the wedge \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.

\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it Eor window}. \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise. In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum. We use an all-sky foreground emission model that includes both diffuse and bright compact components.

@@

For the diffuse component, we use an all-sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all-sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.

A low resolution version of the diffuse foreground model used is show in figure~\ref{fig:DSM}. Contours of the MWA tile power pattern shown in figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction. In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times smaller.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}
